---
title: "Module_4 Assignment"
author: "Dinesh"
date: "2023-09-23"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```
#Further more we need to specify the objective, constraints, direction, and constants in order to solve the problem.
Objective Function

The Objective function is to Max Z = 420(L1 + L2 + L3) + 360(M1 + M2 + M3) + 300(S1 + S2 + S3)
It can be written as Z = 420L1 + 360M1 + 300S1 + 420L2 + 360M2 + 300S2 + 420L3 + 360M3 + 300S3
    L1 + M1 + S1 <= 750
    L2 + M2 + S2 <= 900
    L3 + M3 + S3 <= 450
    20L1 + 15M1 + 12S1 <= 13000
    20L2 + 15M2 + 12S2 <= 12000
    20L3 + 15M3 + 12S3 <= 5000
    L1 + L2 + L3 <= 900
    M1 + M2 + M3 <= 1200
    S1 + S2 + S3 <= 750
Non Negativity Constraints
    L1, L2, L3, M1, M2, M3, S1, S2, S3 >= 0
The above constraints can be written as below


L1 + M1 + S1 + 0*L2 + 0*M2 + 0*S2 + 0*L3 + 0*M3 + 0*S3 <= 750
0L1 + 0M1 + 0S1 + L2 + M2 + S2 + 0L3 + 0M3 + 0S3 <= 900
0L1 + 0M1 + 0S1 + 0L2 + 0M2 + 0S2 + L3 + M3 + S3 <= 450
20L1 + 15M1 + 12S1 + 0L2 + 0M2 + 0S2 + 0L3 + 0M3 + 0S3 <= 13000
0L1 + 0M1 + 0S1 + 20L2 + 15M2 + 12S2 + 0L3 + 0M3 + 0S3 <= 12000
0L1 + 0M1 + 0S1 + 0L2 + 0M2 + 0S2 + 20L3 + 15M3 + 12S3 <= 5000
L1 + 0M1 + 0S1 + L2 + 0M2 + 0S2 + L3 + 0M3 + 0S3 <= 900
0L1 + M1 + 0S1 + 0L2 + M2 + 0S2 + 0L3 + M3 + 0S3 <= 1200
0L1 + 0M1 + S1 + 0L2 + 0M2 + S2 + 0L3 + 0M3 + S3 <= 750

```{r}
library("lpSolve")
#Defining the Objective Function - f.obj
f.obj <- c(420,360,300,420,360,300,420,360,300)

```
#Defining the Constraints - f.con
```{r}
f.con <- matrix(c(1,1,1,0,0,0,0,0,0,
                  0,0,0,1,1,1,0,0,0,
                  0,0,0,0,0,0,1,1,1,
                  20,15,12,0,0,0,0,0,0,
                  0,0,0,20,15,12,0,0,0,
                  1,0,0,1,0,0,1,0,0,
                  0,1,0,0,1,0,0,1,0,
                  0,0,1,0,0,1,0,0,1), nrow = 9, byrow = T)
```
#Defining the Constraints - f.con
```{r}
f.con <- matrix(c(1,1,1,0,0,0,0,0,0,
0,0,0,1,1,1,0,0,0,
0,0,0,0,0,0,1,1,1,
20,15,12,0,0,0,0,0,0,
0,0,0,20,15,12,0,0,0,
0,0,0,0,0,0,20,15,12,
1,0,0,1,0,0,1,0,0,
0,1,0,0,1,0,0,1,0,
0,0,1,0,0,1,0,0,1), nrow = 9, byrow=T)
#Defining the Direction of the constraints - f.dir
f.dir <- c('<=',
          '<=',
          '<=',
          '<=',
          '<=',
          '<=',
          '<=',
          '<=',
          '<=')

```

#Defining the Constants that is the right hand side values - f.rhs

```{r}
f.rhs <- c(750,900,450,13000,12000,5000,900,1200,750)

```
#Calling the lp function to solve the problem based on the objective function i.e. to maximize the profit

```{r}
lp('max',f.obj,f.con,f.dir,f.rhs)
```
#Calling the lp function again to get the values for the variables defined above

```{r}
lp('max',f.obj,f.con,f.dir,f.rhs)$solution
```


```{r}
```


```{r}
```

