---
title: "DEA Assignment-5_811295667"
author: "DINESH YECHURI"
date: "2023-10-29"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# Using benchmarking libraries for DEA
library(Benchmarking)
if (!require(knitr)) {
library(knitr)
}
```


```{r}
#Create data frame with the provided data
data <- data.frame(
  DMU = c("Facility 1", "Facility 2", "Facility 3", "Facility 4", "Facility 5", "Facility 6"),
  Staff_hours_per_day = c(100, 300, 320, 500, 350, 340),
  Supplies_per_day = c(0.3, 0.6, 1.2, 2, 1.4,0.7),
  Reimbursed_patient_days = c(15000,15000,40000,28000,20000,14000),
  Privately_paid_patient_days = c(3500,20000,11000,42000,25000,15000)
)
```


```{r}
#Extracting the only one colume of the DMU
dmu<- data[1:1]
dmu
```


```{r}
#Print the data into table format 
kable(data, format = "pandoc", caption = "Hope Valley Health Care Association")
```


```{r}
#Input format
input <- matrix(c(100, 300, 320, 500, 350, 340,0.3, 0.6, 1.2, 2, 1.4,0.7),ncol=2)
output <- matrix(c(15000,15000,40000,28000,20000,14000,3500,20000,11000,42000,25000,15000),ncol=2)

colnames(input) <- c("Staff_hours_per_day", "Supplies_per_day")
colnames(output) <-c("Reimbursed_patient_days", "Privately_paid_patient_days")

input
output

```
***. DEA Analysis using FDH***
```{r}
#DEA input or output efficiency measures, peers, lambdas and slacks
dea(input,output,RTS="FDH")
```


```{r}
FDH<-dea(input,output,RTS = "FDH")
FDH
```


```{r}
#Show the list of objects calculated
str(FDH)
```


```{r}
#identify the  Efficient value
eff(FDH)
```


```{r}
#Identify the peers value

peers(FDH)
```


```{r}
#Identify the Lambda value
lambda(FDH)
```

```{r}
#Add the Efficiency, Peers & Lambda values in the table
P_FDH <- cbind(data, FDH$eff, FDH$lambda, FDH$peers)
colnames(P_FDH)<- c(names(data),names(input), names(output),
                    'Efficiency','Lambda1','Lambda2','Lambda3',
                    'Lambda4','Lambda5','Lambda6','Peers')
#Showa the table
P_FDH
```


```{r}
#plot the graph for FDH Assumption
dea.plot(input,output,RTS="FDH",txt = rownames(P_FDH),xlab = 'input',ylab = 'output')
##dea.plot(input,output,RTS="fdh",ORIENTATION="in-out",txt=rownames(P_FDH),xlab = 'input',ylab = 'output')
```

***. DEA Analysis using CRS***
```{r}
#DEA input or output efficiency measures, peers, lambdas and slacks
dea(input,output,RTS="CRS")
CRS<-dea(input,output,RTS = "CRS")
CRS
```

```{r}
#Show the list of objects calculated
str(CRS)
```

```{r}
# CRS : Identifing the Efficiency values
eff(CRS)
```

```{r}
#Identifying the Peers values
peers(CRS)
```

```{r}
#Identifying the Lambda values
lambda(CRS)
```

```{r}
#Add the Efficiency, Peers & Lambda values in the table
P_CRS <- cbind(data,CRS$eff, CRS$lambda)
#Name the columns of the table
colnames(P_CRS)<- c(names(data),names(input), names(output),
                    'Efficiency','Lambda1','Lambda2','Lambda3',
                    'Lambda4','Lambda5','Lambda6')
#Sow the table
P_CRS
#plot the graph for CRS Assumption
dea.plot(input,output,RTS="CRS",txt = rownames(P_CRS))
```

***. DEA Analysis using VRS***
```{r}
#DEA input or output efficiency measures, peers, lambdas and slacks
dea(input,output,RTS="VRS")
VRS<-dea(input,output,RTS = "VRS")
VRS
```

```{r}
#Show the list of objects calculated
str(VRS)
```

```{r}
#identifying the efficienty values
eff(VRS)
```

```{r}
#identifying the peers
peers(VRS)
```

```{r}
#identifying the lambda
lambda(VRS)
```

```{r}
#Add the Efficiency, Peers & Lambda values in the table
P_VRS <- cbind(data, VRS$eff, VRS$lambda)
#Name the columns of the table
colnames(P_VRS)<- c(names(data),names(input), names(output),'Efficiency','Lambda1','Lambda2','Lambda3','Lambda4','Lambda5','Lambda6')
#Sow the table
P_VRS
#plot the graph for CRS Assumption
dea.plot(input,output,RTS="VRS",txt = rownames(P_VRS))
dea.plot(input,output,RTS="VRS",ORIENTATION="in-out",txt=rownames(P_VRS))
```


***. DEA Analysis using IRS***
```{r}
#DEA input or output efficiency measures, peers, lambdas and slacks
dea(input,output,RTS="IRS")
IRS<-dea(input,output,RTS = "IRS")
IRS
```

```{r}
#Show the list of objects calculated
str(IRS)
```

```{r}
#identifying the efficency values
eff(IRS)
```

```{r}
#Identifying the peers
peers(IRS)
```

```{r}
#Identifying the lambda
lambda(IRS)
```

```{r}
#Add the Efficiency, Peers & Lambda values in the table
P_IRS <- cbind(data, IRS$eff, IRS$lambda)
#Name the columns of the table
colnames(P_IRS)<- c(names(data),names(input), names(output),
                    'Efficiency','Lambda1','Lambda2','Lambda3',
                    'Lambda4','Lambda5','Lambda6')
#Sow the table
P_IRS
#plot the graph for the CRS Assumption
dea.plot(input,output,RTS="IRS",txt = rownames(P_IRS))
dea.plot(input,output,RTS="irs",ORIENTATION="in-out",txt=rownames(P_IRS))
```

***. DEA Analysis using the DRS***
```{r}
#DEA input or output efficiency measures, peers, lambdas and slacks
dea(input,output,RTS="DRS")

DRS<-dea(input,output,RTS = "DRS")
DRS
```

```{r}
#Show the list of objects calculated
str(DRS)
```

```{r}
#Identifying the Efficiency value
eff(DRS)
```

```{r}
#Identifying the peers
peers(DRS)
```

```{r}
#Identifying the lambda
lambda(DRS)
```

```{r}
#Add the Efficiency, Peers & Lambda values in the table
P_DRS <- cbind(data, DRS$eff, DRS$lambda)
#Name the columns of the table
colnames(P_DRS)<- c(names(data),names(input), names(output),
                    'Efficiency','Lambda1','Lambda2','Lambda3',
                    'Lambda4','Lambda5','Lambda6')
#Sow the table
P_DRS
#plot the graph for CRS Assumption
dea.plot(input,output,RTS="DRS",txt = rownames(P_DRS))
dea.plot(input,output,RTS="Drs",ORIENTATION="in-out",txt=rownames(P_DRS))
```

***. DEA Analysis using FRH***
```{r}
#DEA input or output efficiency measures, peers, lambdas and slacks
dea(input,output,RTS="ADD")
FRH<-dea(input,output,RTS = "ADD")
FRH
```

```{r}
#Show the list of objects calculated
str(FRH)
```

```{r}
#Idendtying the Efficiency values
eff(FRH)
```

```{r}
#Identifying the peers
peers(FRH)
```

```{r}
#Identifying the lambda
lambda(FRH)
```

```{r}
#Add the Efficiency, Peers & Lambda values in the table
P_FRH <- cbind(data, FRH$eff, FRH$lambda)
#Name the columns of the table
colnames(P_FRH)<- c(names(data),names(input), names(output),
                    'Efficiency','Lambda1','Lambda2','Lambda3',
                    'Lambda4','Lambda5','Lambda6')
#Sow the table
P_FRH
#plot the graph for CRS Assumption
dea.plot(input,output,RTS="ADD",txt = rownames(P_FRH))
dea.plot(input,output,RTS="ADD",ORIENTATION="in-out",txt=rownames(P_FRH))

```

## Compare and Contrast the above Results
```{r}
#Add the Efficiency of all the DMUs for all the Assumptions in a table
EfficiencyReport <- cbind(data[,1],FDH$eff,CRS$eff,VRS$eff,IRS$eff,DRS$eff,FRH$eff)
#Name the columns of the table
colnames(EfficiencyReport) <- c(names(dmu),'FDH Efficiency',
                                'CRS Efficiency','VRS Efficiency','IRS Efficiency',
                                'DRS Efficiency','FRH Efficiency')
#Show the Efficiency table
EfficiencyReport
#Let's compare the Efficiency of all the DMUs for all the assumptions using a plot
#Concatenate the Efficiency
spreadsheet <- cbind(FDH$eff,CRS$eff,VRS$eff,IRS$eff,DRS$eff,FRH$eff) 
#Name the rows
rownames(spreadsheet) <- c("Facility1","Facility2","Facility3","Facility4",
                           "Facility5","Facility6")
#Name the columns
colnames(spreadsheet) <- c ("FDH","CRS","VRS","IRS","DRS","FRH")
#See the result
spreadsheet
#plot the graph
barplot(t(spreadsheet),col=gray.colors(6),xlab = "DMUs", ylab="Efficiency",beside=TRUE)
```


Hence,we can conclude that

1) Facility 1 is fully efficient for all the assumptions.

2) Facility 2 is fully efficient for all the assumptions.

3) Facility 3 is fully efficient for FDH, VRS, DRS and FRH assumptions. For assumptions IRS and CRS, it is 87.9% efficient.

4) Facility 4 is fully efficient for all the assumptions.

5) Facility 5 is fully efficient for FDH and FRH assumptions. For assumptions and CRS and DRS it is 89.4% efficient. For VRS and IRS assumption it is 92.3% efficient.

6) Facility 6  FDH and FRH assumptions, it is 88.2% efficient. For CRS and dRS assumptions, it is 70.4% efficient. For VRS and IRS Assumption it is 72.7% efficient.

